---
title: "Cross Validation and Bootstrap"
output: 
  pdf_document: 
    latex_engine: xelatex
---
We use the `set.seed()` function in order to set a seed for `R`'s random number generator, so that the results are reproducible.
```{r}
library(ISLR2)
library(boot)
set.seed(1)
```



# Cross validation

The LOOCV estimate can be automatically computed for any generalized linear model   using the `glm()` and `cv.glm()` functions.  If we use `glm()` to fit a model without passing in the `family` argument, then it  performs linear regression, just like the `lm()` function. `glm()` can be used together with `cv.glm()`. The `cv.glm()` function is part of the `boot` library. The `cv.glm()` function produces a list with several components.  The two numbers in the `delta` vector contain the LOOCV estimates of the test error. The latter is bias corrected version. In the case of LOOCV, these two numbers should be very close to each other.
```{r}
glm.fit <- glm(mpg ~ horsepower, data = Auto)
cv.err <- cv.glm(Auto, glm.fit)
cv.err$delta
```

We can repeat this procedure for increasingly complex polynomial fits. We use `for` loop and the vector `cv.error` is created to store the estimates.

```{r}
cv.error <- rep(0, 10)
for (i in 1:10) {
  glm.fit <- glm(mpg ~ poly(horsepower, i), data = Auto)
  cv.error[i] <- cv.glm(Auto, glm.fit)$delta[1]
}
cv.error
```
We see a sharp drop in the estimated test MSE between the linear and quadratic fits, but then no clear improvement from using higher-order polynomials.

The `cv.glm()` function can also be used to implement $k$-fold CV.
```{r}
set.seed(17)
cv.error.10 <- rep(0, 10)
for (i in 1:10) {
  glm.fit <- glm(mpg ~ poly(horsepower, i), data = Auto)
  cv.error.10[i] <- cv.glm(Auto, glm.fit, K = 10)$delta[1]
}
cv.error.10
```
We still see little evidence that using cubic or higher-order polynomial terms leads to lower test error than simply using a quadratic fit. The two numbers associated with `delta` are essentially the same when LOOCV is performed. When we instead perform $k$-fold CV, then the two numbers associated with `delta` differ slightly.

# Bootstrap
The bootstrap approach can be used  to assess the
variability of the coefficient estimates and predictions from a statistical learning method. Here we use the bootstrap approach in order to assess the variability of
the estimates for $\beta_0$ and $\beta_1$, the intercept and slope terms for the linear regression model
that uses  `horsepower` to predict `mpg` in the `Auto` data set.

We first create a simple function, `boot.fn()`, which takes in the
`Auto` data set as well as a set of indices for the observations, and
returns the intercept and slope estimates for the linear regression model. We then apply this function
to the full set of $392$ observations in order to compute the estimates of $\beta_0$ and $\beta_1$ on the entire data set.
```{r}
boot.fn <- function(data, index)
  coef(lm(mpg ~ horsepower, data = data, subset = index))
boot.fn(Auto, 1:392)
```

Next, we use the `boot()` function to compute the standard errors of 1,000 bootstrap estimates for the intercept and slope terms.
```{r}
set.seed(1)
boot(Auto, boot.fn, 1000)
```

This indicates that the bootstrap estimate for ${\rm SE}(\hat{\beta}_0)$ is $0.84$, and that the bootstrap estimate for ${\rm SE}(\hat{\beta}_1)$ is $0.0073$. We compare them with results from standard formulas.
```{r}
summary(lm(mpg ~ horsepower, data = Auto))$coef
```


The standard error estimates using the standard formulas are $0.717$ for the intercept and $0.0064$ for the slope. Interestingly, these are somewhat different from the
estimates obtained using the bootstrap. Indeed, bootstrap standard errors are automatically heteroskedasticity-robust.


To get White heteroskedasticity robust standard errors, use package `sandwich`. Construct the robust variance covariance matrix `rvcv`. The robust standard errors are on the diagonal.
```{r}
library(sandwich)
lm.fit = lm(mpg ~ horsepower, data = Auto)
rvcv=vcovHC(lm.fit,type="HC")
sqrt(diag(rvcv))
```

To get *t*-statistics and *p*-values:
```{r}
library(lmtest)
coeftest(lm.fit, vcov=rvcv)
```

