---
title: 'Lab 10: Double LASSO'
output:
  pdf_document:
    latex_engine: xelatex
  html_document:
    df_print: paged
---

## Monte Carlo simulations

```{r}
library(hdm)
?rlassoEffect
n=100 
R=300 
rho=1
beta1=0
beta2=0.35
```

Write a function for generating data:
```{r}
data_sim<-function(n,beta1,beta2,rho){
  X=matrix(rnorm(n*3),ncol=3)
  X[,2]<-rho*X[,1]+X[,2]
  Y=beta1*X[,1]+beta2*X[,2]+rnorm(n)
  data<-list(Y=Y,X=X)
}
```

Generate data on the main regressor (D), potential controls, and the dependent variable:
```{r}
set.seed(5,sample.kind = "Rejection")
data<-data_sim(n,beta1,beta2,rho)
y=data$Y #dep. variable
Controls=data$X[,-1] # controls
D=data$X[,1] # the main regressor for which the effect is estimated
```

Run double LASSO:
```{r}
Effect<-rlassoEffect(Controls,y,D,method="double selection")
summary(Effect)
```
Objects inside:
```{r}
names(Effect)
```
Included controls and t-statistic on D:
```{r}
Effect$selection.index
Effect$t
```

We run the simulations using the setup from lab 9.
```{r}
rho=1
set.seed(6064,sample.kind = "Rejection")
T_Beta1_post=rep(0,R) # Vector to store t-stats for the main regressor
for (r in 1:R){
  data<-data_sim(n,beta1,beta2,rho)
  Effect<-rlassoEffect(data$X[,-1],data$Y,data$X[,1],method="double selection")
  T_Beta1_post[r]=Effect$t 
}
```

Plot of the distribution of the post-double-Lasso $t$-statistic:
```{r}
low=min(T_Beta1_post)
high=max(T_Beta1_post)
step=(high-low)/20
hist(T_Beta1_post,breaks=seq(low-2*step,high+2*step,step),xlab="estimates",main="The exact distribution of the post-Double-LASSO t-statistic vs N(0,1)",freq=FALSE,ylim=c(0,0.5))

# add a vertical line at the true value
abline(v=beta1,col="blue")

# add the plot of the N(0,1) pdf 
x=seq(-4,4,0.01)
f=exp(-x^2/2)/sqrt(2*pi) 
lines(x,f,col="red")
```

## Illustration of double LASSO with cross country growth data

The model is $\Delta\log (GDP_{it})=\alpha \cdot GDP_{i0}+U_i$. Hypothesis: $\alpha <0$. Less developed countries catch up with more developed. 
```{r}
data("GrowthData")
?GrowthData
names(GrowthData)
```

The hypothesis fails:
```{r}
summary(lm(Outcome~gdpsh465,data=GrowthData))
```

An alternative model controls for the institutional and technological characteristics: $\Delta\log (GDP_{it})=\alpha \cdot GDP_{i0}+X_i'\beta+U_i$.

There are a lot of potential controls:
```{r}
dim(GrowthData)
```

Let's set up estimation
```{r}
names(GrowthData)
y=as.vector(GrowthData$Outcome)
D=as.vector(GrowthData$gdpsh465)
Controls=as.matrix(GrowthData)[,-c(1,2,3)]
```

We run OLS with all controls. The estimate is negative but the standard error is too large, since there are too many controls.
```{r}
Full=lm(y~D+Controls)
head(coef(summary(Full)),2)
```

Post-LASSO with Double LASSO
```{r}
Effect<-rlassoEffect(Controls,y,D,method="double selection")
summary(Effect)
```

Included controls:
```{r}
Effect$selection.index
sum(Effect$selection.index==TRUE)
```

## The partialling out approach
```{r}
Effect_PO<-rlassoEffect(Controls,y,D,method="partialling out")
summary(Effect_PO)
```

```{r}
Effect_PO$selection.index
sum(Effect_PO$selection.index==TRUE)
```
