---
title: 'Lab 9: Naive Post-LASSO'
output:
  pdf_document:
    latex_engine: xelatex
  html_document:
    df_print: paged
---

In the Monte Carlo simulations below, we illustrate the bias of the naive post-Lasso estimator. The bias comes from the failure of Lasso to reliably detect small non-zero coefficients.

```{r}
library(glmnet)
n=100 #sample size
R=300 #number of Monte Carlo repetitions
```

The data generating process for simulations:
$$\begin{aligned}
Y_i &=\beta_1 X_{i,1}+\beta_2 X_{i,2}+\beta_3 X_{i,3}+U_i,\\
\beta_1 &=0,\\
\beta_2 &=0.35,\\
\beta_3 &=0,\\
X_{i,2} &= \rho X_{i,1}+Z_{i,2},\\
X_{i,1},U_i,Z_{i,2},X_{i,3} &\sim \; \text{iid}\;N(0,1)\;\text{and independent from each other}.
\end{aligned}$$

Three potential regressors:

* Regressor 1 is the main regressor.
* Regressor 2 has a small coefficient and its correlation with Regressor 1 depends on the magnitude of $\rho$.
* Regressor 3 is irrelevant.

```{r}
beta1=0
beta2=0.35
```

## "Large" $\rho$: $\rho=1$

We assume that Regressor 1 is strongly correlated with controls.
```{r}
rho=1
```

We write a function for generating data:
```{r}
data_sim<-function(n,beta1,beta2,rho){
  X=matrix(rnorm(n*3),ncol=3)
  X[,2]<-rho*X[,1]+X[,2]
  Y=beta1*X[,1]+beta2*X[,2]+rnorm(n)
  data<-list(Y=Y,X=X)
}
```

We'll use LASSO with cross validation to select the controls, and then estimate the effect of Regressor 1 on $Y$. We set the penalty weight of Regressor 1 to 0 to always include it.
```{r}
w=rep(1,3)
w[1]=0
data<-data_sim(n,beta1,beta2,rho)
CV.Lasso=cv.glmnet(data$X,data$Y,family="gaussian",alpha=1,penalty.factor=w)
Included=which(coef(CV.Lasso,s=CV.Lasso$lambda.1se)[-1]!=0)
Included
```

We generate data, select covariates using LASSO (with Regressor 1 being always in) and store the t-statistic for the coefficient of the first regressor.
```{r}
rho=1
set.seed(42,sample.kind = "Rejection")
IN2=0 # counter for inclusion of X2
T_Beta1_post=rep(0,R) # Vector to store T-stats for the main regressor
for (r in 1:R){
  data<-data_sim(n,beta1,beta2,rho)
  
  CV.Lasso=cv.glmnet(data$X,data$Y,family="gaussian",alpha=1,penalty.factor=w)
  Included=which(coef(CV.Lasso,s=CV.Lasso$lambda.1se)[-1]!=0)
  
  Post_OLS=lm(data$Y~data$X[,Included])
  T_Beta1_post[r]=coef(summary(Post_OLS))[2,3] #Selects the t-statistic on X1
  
  IN2=IN2+(coef(CV.Lasso,s=CV.Lasso$lambda.1se)[3]!=0)
}
print("Prob. of X2 included")
IN2/R

```

We plot the histogram of the post-LASSO t-statistic for the first regressor. Its asymptotic distribution should be centered around zero since the true coefficient is zero. We compare it with the $N(0,1)$ distribution. Because Regressor 2 is omitted with high probability and correlated with Regressor 1, the distribution is distorted.

```{r}
low=min(T_Beta1_post)
high=max(T_Beta1_post)
step=(high-low)/20
hist(T_Beta1_post,breaks=seq(low-2*step,high+2*step,step),xlab="estimates",main="The exact distribution of the post-LASSO t-statistic vs N(0,1)",freq=FALSE,ylim=c(0,0.5))

# add a vertical line at the true value
abline(v=beta1,col="blue")

# add the plot of the N(0,1) pdf 
x=seq(-4,4,0.01)
f=exp(-x^2/2)/sqrt(2*pi) 
lines(x,f,col="red")
```


## Repeat with a "small" $\rho$: $\rho=0.1$ 

We rerun the simulations with a smaller $\rho$. Regressor 2 is still dropped with a high probability.

```{r}
rho=0.25
set.seed(42,sample.kind = "Rejection")
IN2=0 # counter for inclusion of X2
T_Beta1_post=rep(0,R) # Vector to store T-stats for the main regressor
for (r in 1:R){
  data<-data_sim(n,beta1,beta2,rho)
  
  CV.Lasso=cv.glmnet(data$X,data$Y,family="gaussian",alpha=1,penalty.factor=w)
  Included=which(coef(CV.Lasso,s=CV.Lasso$lambda.1se)[-1]!=0)
  
  Post_OLS=lm(data$Y~data$X[,Included])
  T_Beta1_post[r]=coef(summary(Post_OLS))[2,3] #Selects the t-statistic on X1
  
  IN2=IN2+(coef(CV.Lasso,s=CV.Lasso$lambda.1se)[3]!=0)
}
print("Prob. of X2 included")
IN2/R

```

The exact distribution is less skewed. When the second regressor is only weakly correlated with Regressor 1, there is less distortion.

```{r}
low=min(T_Beta1_post)
high=max(T_Beta1_post)
step=(high-low)/20
hist(T_Beta1_post,breaks=seq(low-2*step,high+2*step,step),xlab="estimates",main="The exact distribution of the post-LASSO t-statistic vs N(0,1)",freq=FALSE,ylim=c(0,0.5))

# add a vertical line at the true value
abline(v=beta1,col="blue")

# add the plot of the N(0,1) pdf 
x=seq(-4,4,0.01)
f=exp(-x^2/2)/sqrt(2*pi) 
lines(x,f,col="red")
```
